### 2015-12-09 paper/paper.tex
Almost complete

\documentclass{article}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{mathrsfs}

\newcommand{\hmmURL}{https://hackage.haskell.org/package/hmm-0.2.1.1/docs/Data-HMM.html}

\title{\bf An HMM-based part-of-speech tagger for Turkish}
\author{Ayberk Tosun\\\texttt{tosun2@illinois.edu}}
\date{}

\begin{document}
%include agda.fmt
\maketitle

\section{Motivation}
\label{sec:motivation}
My main motivation for undertaking this project is to understand the
principles underlying POS-tagging. Due to time and resource constraints, it is
not possible for me to build a state-of-the-art POS-tagger for
Turkish. Nevertheless, I expect my POS-tagger
to work with a decent accuracy that is adequate to be a starting point for
further development. Moreover, \citet{Korkut2015} began the development of a
Turkish NLP library named ``Guguk'', implemented in the programming language Haskell. I intend
to merge this project to Guguk, hence contributing towards
the existence of a Turkish NLP-library in Haskell.

Before proceeding, a few remarks on notational conventions: I use
\texttt{HMM} (with the fixed-width font) to the datatype I'm using from the
\texttt{Data.HMM} package and HMM (with regular font) to refer to the concept.

\section{Related Work}
\label{sec:related_work}

% Briefly survey the most salient prior work that relates to the problem you wish
% to solve. This section should cite relevant sources. This section should be at
% least one to two paragraphs in length.
\cite{dincer2008suffix} built a POS tagger for Turkish using HMMs on suffixes.
This proves an effective approach seeing that (1) Turkish morphology is
exclusively based on suffixes (not prefixes) (2) morphology is central to the
Turkish language so it is likely that any given word is formed through the
agglutination of at least one morpheme. The tagger described has ``\%90.2
accuracy in the best case''.

The widely cited \citet{oflazer1994tagging} outlines the implementation of a
tool for both morpheme glossing and POS tagging, that has ``98-99 \%'' accuracy.
This tool is based on a morphological specification of Turkish and makes use of
a morphological parser called ``PC-Kimmo'' \citep{antworth1991pc}. I should
stress that \citet{oflazer1994tagging} does POS tagging as a by-product of its
capability of morphological parsing.

In general, most of the literature relating to the computational processing of
Turkish pertain primarily to morphology. Compared to the complexity of Turkish
morphology, its syntax seems to be a less interesting problem.

\section{Proposed Work}

I will implement a part-of-speech tagger for the Turkish language. As mentioned
in section~\ref{sec:related_work}, I will mostly focus on parts-of-speech of
words\footnote{I base my definition of a ``word'' on the orthographic space in
  most of the cases. There are certain exceptions to this, such as duplication
  which I denote with the POS tag ``Dup''} and will not be accounting for
morphology, unlike most of the related work I've mentioned. There are definite
problems with this approach. Fixing these would go beyond the scope of this
paper therefore I defer the integration of morphology for further development.

For programming the POS tagger I will be using Haskell, a statically-typed and
purely functional programming language, along with the
\href{\hmmURL}{\texttt{Data.HMM}} package. The data, that is probably the most
important component of this project, comes from the METU-Sabanc{\i} treebank
built by \citet{oflazer2003building}. I am planning to strip the POS tag for
each word from the treebank and use those to train the HMM

\section{Results}
The resulting program I have implemented is based on bigram hidden markov
models. To summarize it we are making use of three probabilities: (1) $P(t_i |
t_{i-1})$ the probability of being at a state (i.e., a tag) given the previous
state. This is computed\footnote{Although I use the word ``compute'', everything
  is already pre-computed and the referred computation is merely a lookup. This
  is discussed in \ref{par:tag}} by the function \texttt{transFn}. (2) $P(w_i | t_i)$ the
probability of seeing a word having a given tag, computed by the function
\texttt{outFn}. Finally, the probability of a given state being the start state
of a sentence, computed by the function \texttt{outProbs}.

\subsection{A Brief Overview of the Program}
The project consists of two executables: (1) \texttt{create-model} that reads
the data from the corpus and parses it into a form from which we can easily
construct an \texttt{HMM}. Unfortunately, it is highly impractical---if not
impossible---to directly serialize \texttt{HMM} since it has three fields
containing functions
\footnote{http://stackoverflow.com/questions/17785916/can-haskell-functions
  -be-serialized for further information.}. We consider this a fault in the
design of the \texttt{Data.HMM} package and urge further development. Although
this has almost no effect on performance, it obfuscates the code.

Since processing the corpus is a resource-demanding operation,
\texttt{create-model} is intended be run only once after which \texttt{run-tag}
can be used to make use of that data.

\subsubsection{\texttt{create-model}}
The executable \texttt{create-model} is implemented in two files:
\texttt{Parse.hs} and \texttt{Main.hs}. \texttt{Parse.hs} handles the parsing of
the corpus.

The function \texttt{freqMap} takes as input a list of orderable elements
(denoted with \texttt{Ord a}) and creates a \texttt{Map} containing, as value,
the frequency of each element.

\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
freqMap :: Ord a $\Rightarrow$ [a] $\rightarrow$ M.Map a Int
freqMap elems = populate empty elems
  where populate :: Ord a Map a Int $\rightarrow$ [a] $\rightarrow$ M.Map a Int
         populate m [] = m
         populate m (x:xs) = let freq = (M.findWithDefault 0 x m) :: Int
                            in populate (M.insert x (freq + 1) m) xs
\end{lstlisting}

This function is called several times with lists of values we want to
\emph{count} such as tag bigrams, a given word with a certain tag et cetera.
Pre-computing the counts is almost mandatory since frequencies (i.e., counts)
will need to be accessed frequently when computing the probabilities. Once the
necessary maps are generated, they are written to files in the directory
\texttt{model}.

\subsubsection{\texttt{tag}}
\label{par:tag}
The executable \texttt{run-tag} is where the actual tagging happens. The
frequencies we have pre-computed in \texttt{create-model} are passed in to the
probability function which we give below.
\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
probability :: (Ord a, Ord b) $\Rightarrow$ (a, b) $\rightarrow$ Map a Int $\rightarrow$ Map (a, b) ℤ $\rightarrow \mathscr{L}$
probability (x, y) c$_1$ c$_2$ = if xCount == 0 || yCount == 0
                                   then logFloat $\epsilon$
                                   else (logFloat yCount) / (logFloat xCount)
  where yCount = fromIntegral (findWithDefault 0 (x, y) c₂) ∷ Double
        xCount = fromIntegral (findWithDefault 0 x c₁) ∷ Double
\end{lstlisting}

This is a straightforward implementation of probability relying on a given count
map. One thing to note is that it returns $\epsilon$ probability (i.e., the
smalles value a \texttt{logFloat}) if a given element is not found.

\texttt{probability} when combined with \texttt{freqMap} forms an
efficient method for implementing Bayes' law. When we are creating the functions
required by \texttt{HMM} we are storing the map returned by a call to
\texttt{freqMap} in the lexical closure of \texttt{probability} by making use
Haskell's currying so that we can eloquently define the two crucial functions as
follows:
\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
transFn s$_1$ s$_2$ = probability (s$_1$, s$_2$) tagFreqs tagBigramFreqs
outFn s e = probability (e, s) wordFreqs taggedWordFreqs
\end{lstlisting}
The probability functions used to create \texttt{transFn} and \texttt{outFn} hold
the count maps \texttt{tagBigramFreqs} and \texttt{taggedWordFreqs} in their
lexical closure, respectively.

Finally we declare a new \texttt{HMM} as follows:

\newpage
\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
      newHMM = HMM { states      = possibleTags ∷ [POS]
                     , events      = ws ∷ [String]
                     , initProbs   = initProbFn
                     , transMatrix = transFn
                     , outMatrix   = outFn
                     }
\end{lstlisting}

This is a simple record (a record is like C struct in Haskell) containing the
probability functions we defined.\footnote{For some reason, the author of the
  \texttt{Data.HMM} library named the fields (i.e., attribute), that are of
  function types, using names that end with \texttt{-Matrix} although they are
  \emph{not} matrices. The reader should not be confused by this.}

\subsection{Performance}
Parsing the entire corpus and creating a model out of it takes 10.5 seconds on
average. Tagging a sentence (which we are assuming to have 5 words on average)
takes $~0.3$ seconds. The implementation of the \texttt{Data.HMM} datatype uses
\texttt{Data.Array}
\footnote{https://hackage.haskell.org/package/array-0.5.1.0/docs/Data-Array.html}
which is why it runs reasonably fast.

\section{Conclusion}

We have shown that it is possible---though not ideal---to build a POS tagger for
Turkish that is based entirely on a bigram hidden markov model. Let us first
look at the accuracy of this model and propose possible fixes that do not
fundamentally change the structure of the model.

\subsection{Accuracy}
\label{subsec:accuracy}

We have tested the POS tagger with 6075 words from a small portion of the corpus
that I reserved for testing (i.e., did not use for tranining the HMM). The POS
tagger tagged 4514 of these words correctly (74.3 \%). Here is a brief list of
known problems:

\begin{description}

  \item[Duplications:] The Turkish language contains many duplications.
  Syntactically speaking, a duplication behaves like a word itself. In the
  treebank, duplications are denoted with two halves joint the with an
  underscore. Since there is not an easy way to detect whether two sequential
  words are reduplications or not, we are giving duplications in an incorrect
  format to the HMM.  We predict that (based on informal tests with the model)
  that correction of this problem would significantly increase the accuracy.

  \item[Abundance of independent clauses:] It is common in the Turkish language
  to have multiple independent clauses in a given sentence. It is possible that
  this decreases the accuracy of the HMM: instead of a tag $t$ being followed by
  a tag $t'$, we get a lot of $t, <clause>, t'$ which obscures the word order we
  are interested in. Since we may always assign the \textsc{Punc} tag to a comma
  with full certainty, we may simply ignore commas as far as probability is
  concerned and tag them with \textsc{Punc}.

\end{description}

\subsection{Limitations}
There is a fundamental limitation to the POS tagger I built: Turkish morphology
is essential to the overall completeness of Turkish.

In English we entertain the notion of \emph{syntactic correctness}, which is
more or less a formalism of correct ordering of words. A speaker of English that
has synactic competence in the language \citep{bachman1990} may know that a
sentence is well-formed, relying on his knowledge of syntactic rules. However,
for the \textit{words}, knowledge of morphological rules is not adequate to
conclude that a word is made up of correctly ordered morphemes; they have to
rely on their lexical knowledge to verify that the morphological form they
derived/inflected is a valid one. Consider the word ``satisfiableness'' which is
not a valid word according to several dictionaries, although being derived with
valid suffixes. The correct form of this concept is ``satisfiability''.

In Turkish, the concept of word validness is a lot like syntactic correctness;
the speaker may rely their knowledge of morphological rules to \emph{generate}
new words. In fact Turkish has \textbf{recursive} morphology \citep{kabak2001}
so a word can be inflected/derived in into infinitely many different forms. This
is why computational processing of Turkish can never fully escape its
morphology.

\subsection{Further research}

A salient possibility for further research is the incorporation of word endings
into the model using regular expressions. It is trivial to express Turkish
suffixes, that change their form as not to upset the major vowel harmony in a
given word. We may say with probably more than $90\%$ probability that a given
word has a certain POS looking whether its ending satisfies regular expressions.

\bibliographystyle{apalike}
\bibliography{your_bibliography}

\end{document}